Parallel multi-objective optimization using co-evolutionary islands and search space overlapping

### 2016-11-18 hpmoon-fgcs.tex
Changed names and all for FGCS

\author{P. Garc\'ia-S\'anchez$^1$, J. Ortega$^2$, J. Gonz\'alez$^2$, P. A. Castillo$^2$ and J.J. Merelo$^2$}
\ead{pablo.garciasanchez@uca.es, \{jortega, jgonzalez, pacv, jmerelo\}@ugr.es}
\address{
$^1$ Department of Computer Science and Engineering. ESI. University of C\'adiz, Spain\\
$^2$ Department of Computer Architecture and Computer Technology.\\ ETSIIT - CITIC. University of Granada, Spain\\
}

@@

In this paper, a new method to automatically adapt the overlapping
sections of the individuals has been compared to other collaborative
% what sections? Decision space sections? Besides, you are using
% section here and before you were using "subset" - JJ
evolutionary methods. 
Different number of islands, chromosome sizes and problems have been
used. The analysis of the obtained experimental results, by using
different  
metrics, shows that coevolutionary approaches can provide
statistically significant improvements with respect to the base
algorithm, being the method to automatically adapt the overlapping
size the one to obtain the best performance during the same execution
time. Results also show that the relation of the number of  
islands (subpopulations) to the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables.n 
\end{abstract}

@@

In our preliminary work in \citep{Garcia16hpmoonANONYMOUS} we tried to demonstrate that using overlapped sections of the chromosome in a coevolutionary multi-objective island-based algorithm can improve the quality of the solutions in the same computing time when the number of islands increases. We discovered that the most adequate overlapping size may vary depending on a relation between the number of islands and the individual size. Moreover, the experiments were conducted in a mono-processor and synchronous island model. This motivates us to continue this research line, but this time, proposing a new method to automatically adapt the size of the overlapped sections, but also, performing all the experiments in a real cluster with up to 128 nodes, with extra number of islands.

@@

% ¿Se podría mencionar por algún lado el trabajo en algoritmos de
% hormigas paralelos de Antonio? - JJ
In our previous work \citep{Garcia16hpmoonANONYMOUS} we used some of
the aforementioned ideas to compare two co-evolutionary dMOEAs. The first
one divided the chromosome in $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome, while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, coevolution with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivates us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Moreover, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and this time, the experiments have been run on a real
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, we are validating the
previous methods.

@@

\subsection{Adaptive coevolution with overlapping islands (A)} 
As in the previous method, the sections to address are overlapped, but instead using one extra section on each side of the $p$ section ($p+1$ and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$), being $c$ a value depending on the number of islands. As a first approach to automatically calculate this value the results shown in \citep{Garcia16hpmoonANONYMOUS} have been used as a base to obtain $c$. In that work, the overlapped method ($c=1$) obtained better results when the number of islands were $>$8. So, we have used the formula $c=round(0.2*P-1)/2$ to calculate the extra sections to overlap. Therefore, when $P=8$ then $c=0$ (equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so on. Figure \ref{fig:adaptive} explains this method.

@@

\begin{figure}
\centering
\includegraphics[width=6cm]{hypervolume.pdf}
\caption{Hypervolume calculation. Areas from Pareto Front solutions (black points) to a reference point (white point) are added.}
\label{fig:hypervolume}
\end{figure}

@@

The ECJ framework \citep{ECJ} has been used to run the
experiments. Specific operators have been developed as new modules for
ECJ, and they can be downloaded from our GitHub repository under a
LGPL V3 License
%\footnote{\url{https://github.com/hpmoon/hpmoon-islands}}. 
\footnote{\url{https://ANONYMOUS-REPOSITORY}}. 
The
island model has been executed synchronously, using the ECJ Internal
Island Model, in a CentOS 5.4 machine with Intel(R) Xeon(R) CPU E5520
@2.27GHz, 16 GB RAM and Java Version 1.6.0\_16.

@@

%DESCRIPTION OF THE RESULTS
Results show that there is a difference in performance when using 512 or 2048 dimensions. Although in some cases the overlapped methods obtain quality metrics values better or significantly equal than the obtained by the baseline, it is clear that 512 dimensions is not a value high enough to apply these methods. This contradicts the results obtained in our previous work in a shared-memory mono-processor version \citep{Garcia16hpmoonANONYMOUS}, where the overlapping method obtained better results than the baseline in both dimension lengths. This can be explained because the added migration latency between nodes requires more time than the time saved during the crossover and migration.

@@

Finally, as happened in \citep{Garcia16hpmoonANONYMOUS}, results show that all the quality indicators decrease when the number of islands is increased, as the  sizes of the subpopulations are smaller. This is consistent with the claim by Dorronsoro et al. \citep{Dorronsoro13superlinear}, who found that cooperative coevolutionary MOEAs work better on bigger populations (more than 100 individuals).

@@

\section*{Acknowledgements}
ANONYMOUS
%This work has been partially funded by projects TIN2014-56494-C4-3-P y TIN2015-67020-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR),  and  MSTR (PRY142/14, Fundaci\'on P\'ublica Andaluza Centro de Estudios Andaluces en la IX Convocatoria de Proyectos de Investigaci\'on).

@@

\bibliography{hpmoon-hpca}
%\begin{thebibliography}{99}
%\bibitem[Kopka and Daly(2003)]{R1}
%Kopka~H and Daly~PW (2003) \textit{A Guide to \LaTeX}, 4th~edn.
%Addison-Wesley.%

>>>>

\author{P. Garc\'ia-S\'anchez$^1$, J. Ortega$^2$, J. Gonz\'alez$^2$, P. A. Castillo$^2$ and J.J. Merelo$^2$}
\ead{pablo.garciasanchez@uca.es, \{jortega, jesusgonzalez, pacv, jmerelo\}@ugr.es}
\address{
$^1$ Department of Computer Science and Engineering. ESI. University of C\'adiz, Spain\\
$^2$ Department of Computer Architecture and Computer Technology.\\ ETSIIT - CITIC. University of Granada, Spain\\
}

@@

In this paper, a new method to automatically adapt the overlapping
sections of the individuals has been compared to other collaborative
% what sections? Decision space sections? Besides, you are using
% section here and before you were using "subset" - JJ
evolutionary methods. 
Different number of islands, chromosome sizes and problems have been
used. The analysis of the obtained experimental results, by using
different  
metrics, shows that coevolutionary approaches can provide
statistically significant improvements with respect to the base
algorithm, being the method to automatically adapt the overlapping
size the one to obtain the best performance during the same execution
time. Results also show that the relation of the number of  
islands (subpopulations) to the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables.
\end{abstract}

@@

In our preliminary work in \citep{Garcia16hpmoon} we tried to demonstrate that using overlapped sections of the chromosome in a coevolutionary multi-objective island-based algorithm can improve the quality of the solutions in the same computing time when the number of islands increases. We discovered that the most adequate overlapping size may vary depending on a relation between the number of islands and the individual size. Moreover, the experiments were conducted in a mono-processor and synchronous island model. This motivates us to continue this research line, but this time, proposing a new method to automatically adapt the size of the overlapped sections, but also, performing all the experiments in a real cluster with up to 128 nodes, with extra number of islands.

@@

% ¿Se podría mencionar por algún lado el trabajo en algoritmos de
% hormigas paralelos de Antonio? - JJ
In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two co-evolutionary dMOEAs. The first
one divided the chromosome in $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome, while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, coevolution with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivates us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Moreover, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and this time, the experiments have been run on a real
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, we are validating the
previous methods.

@@

\subsection{Adaptive coevolution with overlapping islands (A)} 
As in the previous method, the sections to address are overlapped, but instead using one extra section on each side of the $p$ section ($p+1$ and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$), being $c$ a value depending on the number of islands. As a first approach to automatically calculate this value the results shown in \citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that work, the overlapped method ($c=1$) obtained better results when the number of islands were $>$8. So, we have used the formula $c=round(0.2*P-1)/2$ to calculate the extra sections to overlap. Therefore, when $P=8$ then $c=0$ (equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so on. Figure \ref{fig:adaptive} explains this method.

@@

\begin{figure}
\centering
\includegraphics[width=6cm]{hypervolume.jpg}
\caption{Hypervolume calculation. Areas from Pareto Front solutions (black points) to a reference point (white point) are added.}
\label{fig:hypervolume}
\end{figure}

@@

The ECJ framework \citep{ECJ} has been used to run the
experiments. Specific operators have been developed as new modules for
ECJ, and they can be downloaded from our GitHub repository under a
LGPL V3 License
\footnote{\url{https://github.com/hpmoon/hpmoon-islands}}. 
%\footnote{\url{https://ANONYMOUS-REPOSITORY}}. 
The
island model has been executed synchronously, using the ECJ Internal
Island Model, in a CentOS 5.4 machine with Intel(R) Xeon(R) CPU E5520
@2.27GHz, 16 GB RAM and Java Version 1.6.0\_16.

@@

%DESCRIPTION OF THE RESULTS
Results show that there is a difference in performance when using 512 or 2048 dimensions. Although in some cases the overlapped methods obtain quality metrics values better or significantly equal than the obtained by the baseline, it is clear that 512 dimensions is not a value high enough to apply these methods. This contradicts the results obtained in our previous work in a shared-memory mono-processor version \citep{Garcia16hpmoon}, where the overlapping method obtained better results than the baseline in both dimension lengths. This can be explained because the added migration latency between nodes requires more time than the time saved during the crossover and migration.

@@

Finally, as happened in \citep{Garcia16hpmoon}, results show that all the quality indicators decrease when the number of islands is increased, as the  sizes of the subpopulations are smaller. This is consistent with the claim by Dorronsoro et al. \citep{Dorronsoro13superlinear}, who found that cooperative coevolutionary MOEAs work better on bigger populations (more than 100 individuals).

@@

\section*{Acknowledgements}
This work has been partially funded by projects TIN2014-56494-C4-3-P y TIN2015-67020-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR),  and  MSTR (PRY142/14, Fundaci\'on P\'ublica Andaluza Centro de Estudios Andaluces en la IX Convocatoria de Proyectos de Investigaci\'on).

@@

\bibliography{hpmoon-fgcs}
%\begin{thebibliography}{99}
%\bibitem[Kopka and Daly(2003)]{R1}
%Kopka~H and Daly~PW (2003) \textit{A Guide to \LaTeX}, 4th~edn.
%Addison-Wesley.%
